---
title: "Homework2"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##ANLY 530 Assignment 2

Introduction to statistical learning, chapter 2, Exercise - 8 


```{r }

getwd()
setwd("C:/Users/Madhu/Dohaa Tahha/ML1")
getwd()

```

Check all the files in the working directory

```{r}

dir()

```

Read in dataset
```{r}

college =read.csv("College.csv")

```
Structure of the data 

```{r}
str(college)

```


```

Separating the first row that lists the college names from the rest of the data 

```{r}
rownames(college) = college[,1]
fix(college)
college = college[,-1]
fix(college)

```

```{r}
summary(college)

```

Creating plots for the first 10 variables

```{r}
pairs(college[,1:10])

```

Creating box plot for the variables "Outstate" versus "Private"

```{r}
plot(college$Outstate~college$Private)
```

Creating new variable to find out how many colleges are in the elite group

```{r}
Elite = rep("NO", nrow(college))
Elite[college$Top10perc>50]= "Yes"
Elite = as.factor(Elite)
college = data.frame(college,Elite)

```

Using Summary to see if Elite colleges ahve been created in a separate column

```{r}
summary(college)
```

Creating box plot for "Outstate" vs "Elite"

```{r}
plot(college$Outstate~college$Elite)

```

Plotting Histograms for variables namely "Apps", "Enroll", "Accept" , "Outstate"

```{r}
hist(college$Apps,xlim=c(0,20000))
hist(college$Accept,xlim=c(0,10000))
hist(college$Enroll,xlim=c(0,5000))
hist(college$Outstate)

```