---
title: "SVM&Naive_Bayes"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Step 1: collecting data 

```{r }
getwd()
setwd('C:/Users/Madhu/Side Projects/Machine_Learning_Course')

#Read data file

news <- read.csv('OnlineNewsPopularity.csv')

str(news)

```

Step 2: Exploring the data


```{r }
#Using sumary to check the statistics of shares column in the dataset

summary(news$shares)

# Adding new variable Popularity based on the market share value. Considering the median value of 1400 to make a splut in the data as popular and unpopular 

news$Popularity <- ifelse( news$shares <= 1400, "Unpopular","Popular")

news$Popularity <- as.factor(news$Popularity)


#Using table to see how many were defaulted and how many were not

table(news$Popularity)

```


##Method1: Support vector machines

Creating the training and testing data set

```{r }

set.seed(12345)

news_rand <- news[order(runif(39644)),]


# Choosing 75% for training set and remaining 

news_train <- news_rand[1:29733,]
news_test <- news_rand[29734:39644,]


```

 Training model on data 


```{r }
# Installing  kernlab

#install.packages('kernlab')


library(kernlab)

news_classifier <- ksvm(Popularity ~ ., data = news_train[c(29:60,62)], kernel = "vanilladot")

news_classifier

```

There us a training error of 40%

The last and final step 4 of evaluating model performance

```{r }

news_predictions <- predict(news_classifier, news_test)

table(news_predictions, news_test$Popularity)

agreement <- news_predictions == news_test$Popularity

table(agreement)

```

 Model has predicted 5942 correctly and 3970 was misclassified. 


#Method 2: Naive Bayes Algorithm  


Training a model on the data

```{r }

#install.packages('naivebayes')

library(naivebayes)

naive <- naive_bayes(Popularity ~ ., data = news_train[c(29:60,62)])


naive

```


#step 4: Evaluating Model performance

```{r }

naive_predict <- table(predict(naive, news_test), news_test$Popularity)
naive_predict

```

```{r }

#calculate accuracy of the model 

accuracy <- sum(diag(naive_predict))/sum(naive_predict)*100
accuracy

```

```{r}
 prediction <- predict(naive, news_test)
 check <- prediction == news_test$Popularity
 table(check)

```

As seen, the two evaulation methods match.  
